NumPy---Array Computation for Python

### 2019-12-10 paper.tex
Add intro and history first draft

>>>>

NumPy is the fundamental array computation library for the Python
ecosystem.  It underpins almost every library that does numerical
computation, including SciPy.  NumPy consists of the `ndarray`-object
along with utility functions to operate on such arrays.  Because of
its inherent simplicity—being a pointer to memory with some
associated meta-data about shape, data-type, and so forth—the NumPy
array has become the {\it de facto} exchange format for array data in
Python.  That said, NumPy has now seen such widespread adoption that
not only the array object but also its {\it Application Programming
  Interface} (API) has become ubiquitous, and is imitated by, e.g.,
popular tensor computation libraries.

Over the past few years, with external financial support, NumPy has
seen renewed developer interest.  It is actively developed to meet the
needs of its community, including the addition of some fairly large
changes as outlined in Section XXX.

\section*{History}

NumPy was, at its inception, the unifier of the Python array
processing world.  Before NumPy, two packages—Numeric and
numarray—existed.  The first was developed by Paul DuBois and
colleagues at [the lab], and was aimed at fast manipulation of many
small arrays.  In contrast, to handle large astronomical images,
NumArray was developed [by STSCI?]. In other words, these array
objects had different strengths and weaknesses, and therefore tended
to divide the already small user community.  In 200x, Travis Oliphant
wrote NumPy, a ``best of both worlds'' unification of the two
projects, based on Numeric.  The arrival of stable version 1.0 in 200x unified the scientific Python
community, and heralded the era of popular scientific computing in
Python that has led to formidable successes such as [cite some impacts here].

While, by 2006, NumPy was already eminently usable, it lacked good
documentation. In order to use NumPy in his classes, Joe Harrington
from the University of South Florida sponsored several rounds of the
NumPy Documentation Marathon [citations].  During these intensive
documentation improvement sessions, a Wikipedia-like system was
developed to allow community members to contribute so-called
``docstrings''—text describing each NumPy function. A community
documentation standard was published so that these pieces of
documentation would all provide similar information, and conform to a
basic standard of quality. Experienced community members reviewed and
integrated contributed docstrings, and soon NumPy had thorough
documentation coverage.

For many years, NumPy operated without external funding, development
time mostly being contributed freely by students and researchers in
their spare time.  In-person meetings were sponsored off of
individual's research grants, and while industry sponsored
contributions were made, notably those by Mark Wiebe when he was
sponsored by Travis Oliphant at X, mostly the package had rapid growth
and adoption without external investment.  In 2014, NumPy received its
first large grants from the Moore and Sloan foundations, totaling 1.3M
dollars.  This allowed for hiring several full time programmers and,
with the support of community members, mobilized a project that had
become stagnant for the several years prior.

NumPy is now actively developed, with weekly online developer
meetings, a formalized NumPy enhancement proposal process, and
multiple active grants (but still almost no industry support).

For a more detailed history of the scientific Python ecosystem, please
refer to [SciPy arxiv / nature paper].
